The benefits of prefetching for large-scale cloud-based neuroimaging analysis workflows

### 2021-10-05 paper/paper.tex
cut some words

\subsection{Experiments}
To evaluate the performance Rolling Prefetch, we conducted 4 experiments: 1) varying
the number of files; 2) varying the blocksize; 3) parallel processing of files;  and 4) neuroimaging use-cases.
For experiments 1-3, the pipeline consisted of reading the file with Nibabel. As Nibabel
performs a minimal amount of computing, Rolling Prefetch is believed to have an effect even here. Additional compute required
by real analysis should only increase the benefit of using prefetching.
% For all experiments we are using tractography
% files belonging to the HYDI dataset stored on S3.

@@

We chose not to modify the pipeline, in order to determine what is the speedup provided
by Rolling Prefetch without additional optimization, and in order to be able to speculate on the possible benefits of adapting existing pipelines
to support larger-than-memory files and allowing reads and computations to occur together.

@@

At a very large number of blocks, the performance of both \sfs and Rolling Prefetch is degraded due to
increased impacts of both S3 storage latency and additionally, in the case of Rolling Prefetch, local storage latency (Figure~\ref{fig:blocksize}).  \sfs starts to surpass the speed of Rolling Prefetch, although not substantially (approx. 1.03~X) at 748 blocks.  Furthermore, while we read from S3
at certain blocksizes, the block size on local storage is dictated by filesystem and kernel readahead 
size. In our specific case, the data would be read from tmpfs, and thus we do not see a significant added cost
to Rolling Prefetch. It could be inferred that the overhead of Rolling Prefetch may be more significant with a large
number of blocks due to Nibabel's access patterns, which consist of many small reads.

Conversely, when decreasing the number of blocks we start to see an increase in performance in both \sfs and Rolling Prefetch. This
is because both benefit from decreased latency to access data on S3, by generally making fewer read calls to it. Rolling Prefetch is found to be faster than \sfs with a peak speedup of 1.24~X at 187 blocks (i.e. 
\SI{32}{\mebi\byte} blocks). The speedup obtained from Rolling Prefetch did not vary significantly between 24 and
187 blocks, averaging at a speedup of approximately 1.2~X. Rolling Prefetch outpaces \sfs at less than
748 blocks as the cost of local latency has diminished sufficiently with the larger blocks. As the blocksize
increases, Rolling Prefetch is able to begin masking S3 latency and bandwidth within its compute.
% The speedup being
% proportional to how many more blocks can be read within the computation.
\sfs and Rolling Prefetch performance converges again when the blocks reach a certain size, placing more weight disk bandwidth
than latency. \sfs exceeds the performance of Rolling Prefetch at a single block, where no actual prefetching can take
place and we pay the cost of implementation overheads, such as writing the data to local storage 
and reading it from local storage rather than directly from memory.

@@

Rolling Prefetch is an added layer on top of \sfs that replaces its built-in cache strategies to intercept
application read calls and ensure that data is preloaded. The implementation ensures that filesystem space requirements are not exceeded during prefetching through limits set by the user. 
With the built-in eviction mechanism, users are not required to do any form of data management should they be processing datasets larger than available storage. Furthermore, the library allows configuration of multiple storage device as cache for prefetching. Currently,
files are written to specific storage devices based on available space and storage priority in the list.

With just a simple implementation performed on a computation based on reading, we have observed up to a 1.8$\times$ speedup and a maximum overhead of 1.03$\times$. These observed speedups were
obtained when we set the cache location to a \texttt{tmpfs} device and may naturally decrease with a slower device such as an SSD. In our particular case, the speedups meant saving 20~min of processing time
on loading nearly \SI{100}{\gibi\byte} of data with a maximum runtime of approximately 71~min. Moreover, this
was achieved on a small instance with only \SI{1}{\gibi\byte} of prefetch storage allotted to each parallel
process, indicating that we can achieve good speedups even when resources are constrained.

@@

The risk with passing large amounts data to the library within a task is that it is less robust to failures as any failed task would have to resume from the beginning. This could
take a significant amount of time, depending on how much data had been processed by the task prior to failure.
% Making the tasks smaller would reduce the costs associated with task restart, but then the benefits of
% Rolling Prefetch would be limited.
It is understandable that implementations such as Netco are adaptations to
persistent filesystem, since then prefetching could work without limiting scheduler logic. For our 
implementation to be efficient within parallel processing frameworks, we would need to decouple it from
the task itself and allow it to communicate with the task scheduler to determine which files to prefetch and
where.

>>>>

\subsection{Experiments}
To evaluate the performance Rolling Prefetch, we conducted four experiments: 1) varying
the number of files; 2) varying the blocksize; 3) parallel processing of files;  and 4) neuroimaging use-cases.
For experiments 1-3, the pipeline consisted of reading the file with Nibabel. As Nibabel
performs a minimal amount of computing, Rolling Prefetch is believed to have an effect even here. Additional compute required
by real analysis should only increase the benefit of using prefetching.
% For all experiments we are using tractography
% files belonging to the HYDI dataset stored on S3.

@@

We chose not to modify the pipeline in order to determine what is the speedup provided
by Rolling Prefetch without additional optimization, and in order to be able to speculate on the possible benefits of adapting existing pipelines
to support larger-than-memory files and allowing reads and computations to occur together.

@@

At a very large number of blocks, the performance of both \sfs and Rolling Prefetch is degraded due to
increased impacts of both S3 storage latency and, in the case of Rolling Prefetch, local storage latency (Figure~\ref{fig:blocksize}).  \sfs starts to surpass the speed of Rolling Prefetch, although not substantially (approx. 1.03~X) at 748 blocks.  Furthermore, while we read from S3
at certain blocksizes, the block size on local storage is dictated by filesystem and kernel readahead 
size. In our specific case, the data would be read from tmpfs, and thus we do not see a significant added cost
to Rolling Prefetch. It could be inferred that the overhead of Rolling Prefetch may be more significant with a large
number of blocks due to Nibabel's access patterns, which consist of many small reads.

Conversely, when decreasing the number of blocks we see an increase in performance in both \sfs and Rolling Prefetch. This
is because both benefit from decreased latency to access data on S3, by making fewer read calls to it. Rolling Prefetch is found to be faster than \sfs with a peak speedup of 1.24~X at 187 blocks (i.e. 
\SI{32}{\mebi\byte} blocks). The speedup obtained from Rolling Prefetch did not vary significantly between 24 and
187 blocks, averaging at a speedup of approximately 1.2~X. Rolling Prefetch outpaces \sfs at less than
748 blocks as the cost of local latency has diminished sufficiently with the larger blocks. As the blocksize
increases, Rolling Prefetch is able to begin masking S3 latency and bandwidth within its compute.
% The speedup being
% proportional to how many more blocks can be read within the computation.
\sfs and Rolling Prefetch performance converges again when the blocks reach a certain size, placing more weight disk bandwidth
than latency. \sfs exceeds the performance of Rolling Prefetch at a single block, where no actual prefetching can take
place and we pay the cost of implementation overheads, such as writing the data to local storage 
and reading it from local storage rather than directly from memory.

@@

Rolling Prefetch is an added layer on top of \sfs that replaces its built-in cache strategies to intercept
application read calls and ensure that data is preloaded. The implementation ensures that filesystem space requirements are not exceeded during prefetching through limits set by the user. 
With the built-in eviction mechanism, users are not required to do any form of data management should they be processing datasets larger than available storage. Furthermore, the library allows configuration of multiple storage device as cache for prefetching. Currently,
files are written to specific storage devices based on available space and storage priority.

With just a simple data loading workload, we have observed up to a 1.8$\times$ speedup and a maximum overhead of 1.03$\times$. These observed speedups were
obtained when we set the cache location to a \texttt{tmpfs} device and may naturally decrease with a slower device such as an SSD. In our particular case, the speedups meant saving 20~min of processing time
on loading nearly \SI{100}{\gibi\byte} of data with a maximum runtime of approximately 71~min. Moreover, this
was achieved on a small instance with only \SI{1}{\gibi\byte} of prefetch storage allotted to each parallel
process, indicating that we can achieve good speedups even when resources are constrained.

@@

The risk with passing large amounts data to the library within a task is that it is less robust to failures as any failed task would have to resume from the beginning. This could
take a significant amount of time, depending on how much data had been processed by the task prior to failure.
% Making the tasks smaller would reduce the costs associated with task restart, but then the benefits of
% Rolling Prefetch would be limited.
It is understandable that implementations such as Netco are adaptations to
persistent filesystem, since then prefetching could work without limiting scheduler logic. For our 
implementation to be efficient within parallel processing frameworks, we would need to decouple it from
the task itself and allow it to communicate with the task scheduler to determine which files to prefetch.
